Exception Handling Bug Hazards in Android

### 2016-04-20 SPRINGER_JOURNAL/msr_journal.tex
fixing typos

This analysis revealed that approximately 75\% of the exceptions
 that caused the stack traces are implicitly thrown by the runtime
environment  due to mistakes in the programming logic (e.g., out-of-bounds array index, null pointer
references) and resource limitations. Although such exceptions do not directly point to violations of
the best practices described before (which are related to the
explicitly thrown exceptions) they impose a major threat to app robustness,
and therefore represent a critical bug hazard to the exception
handling code of Android applications.
Security and concurrency, which are known to be critical issues for Android apps,
raised few of the reported exceptions (less than 5\% of the analysed stack traces).

@@

\emph{\textbf{Runtime Exception wrapping a Checked Exception.}}
This wrapping was responsible for 49.5\% of
the cross-type wrappings.  From this set 50\%  were performed by methods defined by the Android platform.
We observe that this is a common implementation practice in the methods of the Android platform.
According to Java best practices checked exceptions represent conditions from which the caller
is expected to recover. By converting a checked exception in a general Runtime class,
besides loosing contextual information about the exception,
the client can simply ignore such exception - which can negatively impact the app robustness.
Although it is considered a bad practice, when such wrapping is accompained with a well defined
exception handling policy (which assures that proper handling will be provided within the app)
the impact on app robustness can be mitigated.
%\textcolor{red}

@@

%\textcolor{red}
Our last research question addresses whether the exception handling code helps with the development of robust Android applications, and what developers usually do to prevent apps from crashing. The motivation behind this question was to discover common practices for dealing with uncaught exceptions and to acess developers opinion about the role of the exception handling mechanism in the development of robust apps. Hence, the last question is as follows:

@@

\emph{Survey Customization.} Although the survey was applied to developers who had contributed to at least one of the GitHub apps, analyzed in the mining study, the survey was not customized to each developer. In other words, it did not contain questions focusing specific bug hazards identified in their apps. That customization of the survey (for each specific developer/app) could give more insights about the exception handling usage and bug hazards, and could even have improved the response rate for the survey. However, we did not followed this approach because the high number of projects analyzed (482 Android apps) and their contributors (1,824 developers) would increase the complexity and time needed (i) to customize the survey and specially (ii) to analyze the survey responses - since themes could emerge from specific contexts. This approach can be very useful in a guided interview involving some of the respondents to refine the findings of this exploratory survey, in a future work.

>>>>

This analysis revealed that approximately 75\% of the exceptions
 that caused the stack traces are implicitly thrown by the runtime
environment  due to mistakes in the programming logic (e.g., out-of-bounds array index, null pointer
references) and resource limitations. Although such exceptions do not directly point to violations of
the best practices described before (which are related to the
explicitly thrown exceptions) they impose a major threat to app robustness,
and therefore represent a critical bug hazard to the exception
handling code of Android applications.
Security and concurrency, which are known to be critical issues for Android apps,
raised few of the reported exceptions (less than 5\% of the analyzed stack traces).

@@

\emph{\textbf{Runtime Exception wrapping a Checked Exception.}}
This wrapping was responsible for 49.5\% of
the cross-type wrappings.  From this set 50\%  were performed by methods defined by the Android platform.
We observe that this is a common implementation practice in the methods of the Android platform.
According to Java best practices checked exceptions represent conditions from which the caller
is expected to recover. By converting a checked exception in a general Runtime class,
besides loosing contextual information about the exception,
the client can simply ignore such exception - which can negatively impact the app robustness.
Although it is considered a bad practice, when such wrapping is accompanied by a well defined
exception handling policy (which assures that proper handling will be provided within the app)
the impact on app robustness can be mitigated.
%\textcolor{red}

@@

%\textcolor{red}
Our last research question addresses whether the exception handling code helps with the development of robust Android applications, and what developers usually do to prevent apps from crashing. The motivation behind this question was to discover common practices for dealing with uncaught exceptions and to assess developers perspective about the role of the exception handling mechanism in the development of robust apps. Hence, the last question is as follows:

@@

\emph{Survey Customization.} Although the survey was applied to developers who had contributed to at least one of the GitHub apps, analyzed in the mining study, the survey was not customized to each developer. In other words, it did not contain questions focusing on specific bug hazards identified in their apps. That customization of the survey (for each specific developer/app) could give more insights about the exception handling usage and bug hazards, and could even have improved the response rate for the survey. However, we did not followed this approach because the high number of projects analyzed (482 Android apps) and their contributors (1,824 developers) would increase the complexity and time needed (i) to prepare the survey and specially (ii) to analyze the survey responses - since themes could emerge from specific contexts. This approach can be very useful in a guided interview involving some of the respondents to refine the findings of this exploratory survey in a future work.
